Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-15 paper/main.tex
A little bit country
Moved captions to top in tables, and minor spell corrections

\paragraph{Cost and Performance.} Cost and Performance are considered to be the
2nd most important criterion. The absolute minimum for performance is that the
SHA-3 candidate should be faster than the functions in the \ac{SHA}-2 family.
Cost is a measure of how much memory an implementation requires in software,
both the implementation itself and the use of \ac{RAM} during runtime. Another
measure of cost is how many logic gates it takes to implement the function in
hardware.

\paragraph{Algorithm and implementation characteristics.} By this 3rd criterion,
\ac{NIST} emphasizes that algorithms with greater flexibility will be given
preference over other algorithms \cite{s_nistround2}. By flexible, they imply
the possibility of the function to run efficiently on a variety of platforms,
and to use parallelism and instruction set extensions. Another key point of a
flexible hash function is that it should have a simple and elegant design to
encourage understanding, analysis and design confidence.

@@

Tables \ref{tbl:tahoe:datauploaded} and \ref{tbl:tahoe:datadownloaded} and
describes the distribution of the different input sizes of data that
Tahoe-\ac{LAFS} sends to the hash function when uploading and downloading the
test vectors. E.g. to upload the 100 1-byte files, 15700 hashes had to be
calculated, and about 64\% of these were of size 106 byte.

For every upload Tahoe-\ac{LAFS} does, it hashes a size close to the original
file size twice. This does not happen when downloading the file.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhileuploading1001Bfiles.pdf}
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhiledownloading1001Bfiles.pdf}
    \caption{Figures for hash operations of 100 1B files.}
    \label{fig:graph:1001b}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhileuploading1001KBfiles.pdf}
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhiledownloading1001KBfiles.pdf}
    \caption{Figures for hash operations of 100 1KB files.}
    \label{fig:graph:1001kb}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhileuploading501MBfiles.pdf}
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhiledownloading501MBfiles.pdf}
    \caption{Figures for hash operations of 50 1MB files.}
    \label{fig:graph:501mb}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhileuploading5100MBfiles.pdf}
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhiledownloading5100MBfiles.pdf}
    \caption{Figures for hash operations of five 100MB files.}
    \label{fig:graph:5100mb}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhileuploading11GBfile.pdf}
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhiledownloading11GBfile.pdf}
    \caption{Figures for hash operations of one 1GB file.}
    \label{fig:graph:11gb}
\end{figure}

@@

\section{Exclusion of mutable file testing}

@@

>>>>

\paragraph{Cost and Performance.} Cost and Performance are considered to be the
second most important criterion. The absolute minimum for performance is that the
SHA-3 candidate should be faster than the functions in the \ac{SHA}-2 family.
Cost is a measure of how much memory an implementation requires in software,
both the implementation itself and the use of \ac{RAM} during runtime. Another
measure of cost is how many logic gates it takes to implement the function in
hardware.

\paragraph{Algorithm and implementation characteristics.} By this third criterion,
\ac{NIST} emphasizes that algorithms with greater flexibility will be given
preference over other algorithms \cite{s_nistround2}. By flexible, they imply
the possibility of the function to run efficiently on a variety of platforms,
and to use parallelism and instruction set extensions. Another key point of a
flexible hash function is that it should have a simple and elegant design to
encourage understanding, analysis and design confidence.

@@

Tables \ref{tbl:tahoe:datauploaded} and \ref{tbl:tahoe:datadownloaded} and
describes the distribution of the different input sizes of data that
Tahoe-\ac{LAFS} sends to the hash function when uploading and downloading the
test vectors. E.g. to upload the 100 1-byte files, 15700 hashes had to be
calculated, and about 64\% of these were of size 106 byte. For every upload
Tahoe-\ac{LAFS} also hashes a size close to the original file size twice. This
does not happen when downloading the file.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhileuploading1001Bfiles.pdf}
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhiledownloading1001Bfiles.pdf}
    \caption{Graph of timing of hash operations for 100 1B files.}
    \label{fig:graph:1001b}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhileuploading1001KBfiles.pdf}
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhiledownloading1001KBfiles.pdf}
    \caption{Graph of timing of hash operations for 100 1KB files.}
    \label{fig:graph:1001kb}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhileuploading501MBfiles.pdf}
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhiledownloading501MBfiles.pdf}
    \caption{Graph of timing of hash operations for 50 1MB files.}
    \label{fig:graph:501mb}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhileuploading5100MBfiles.pdf}
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhiledownloading5100MBfiles.pdf}
    \caption{Graph of timing of hash operations for five 100MB files.}
    \label{fig:graph:5100mb}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhileuploading11GBfile.pdf}
    \includegraphics[width=0.9\columnwidth]
        {graphs/Timespenthashingwhiledownloading11GBfile.pdf}
    \caption{Graph of timing of hash operations for one 1GB file.}
    \label{fig:graph:11gb}
\end{figure}

@@

\section{Exclusion of mutable files}

@@

\subsection{Possible gain for Tahoe-LAFS}

Even though few of the \ac{SHA}-3 implementations used by us managed to
outperform the implementation of double \ac{SHA}-256 in Tahoe-\ac{LAFS}, some
did for large files, which tells us there is a performance gain by switching
cryptographic hash function in Tahoe-\ac{LAFS}. \autoref{tbl:tahoe:gain}
compares the performance gains for BLAKE, which performed the best in our
measurements, versus \ac{SHA}-256, with the removal of overhead caused by our
measurements. We achieved as much as a 12.35\% gain when uploading the 100MB
files. The gain is smaller on downloads than uploads, this is
mostly caused by our Tahoe-\ac{LAFS} configuration, where we need to upload
shares to three servers, while we only need to download two of them to get the
original file.

\input{tables/tahoe_gain}

On another note, most individuals will not have access to the same high
bandwidth network that our Tahoe-\ac{LAFS} servers ran on, 1Gbps, or even
100Mbps. For an individual with a normal Internet connection, the network is the
bottleneck, and therefore the performance increase with a faster hash algorithm
implementation would not decrease the transfer time for large files. If we
consider Tahoe-\ac{LAFS} for organisations and enterprises which typically
empowers high bandwidth networks, the change could indeed be noticeable.

\subsection{Faster SHA-3 implementations will appear}

The observation should also be made that \ac{SHA}-256 has probably been
optimized for the better part of a decade, while the \ac{SHA}-3 candidates are
relatively new. This will probably mean that in the future we will see even
faster versions of the \ac{SHA}-3 candidates, while it is more probable that
\ac{SHA}-2 is nearing the limit of its performance.
